---
title: "Practical 7"
output: html_notebook
---

```{r}
if (!require("lubridate")){install.packages("lubridate")}
library(lubridate)

if (!require("tidyverse")){install.packages("tidyverse")}
library(tidyverse)

if (!require("janitor")){install.packages("janitor")}
library(janitor)

if (!require("haven")){install.packages("haven")}
library(haven)
```

Q1
part i:
```{r}
medrecord<-readRDS("medrecord.RDS")
str(medrecord)

# Convert date_update to proper Date format
medrecord$date_update<-as.Date(substr(medrecord$date_update,5,nchar(medrecord$date_update)),format="%d/%m/%Y")
head(medrecord$date_update)
summary(diagnoses$date_diag)

# Convert region to factor
medrecord$region<-as.factor(medrecord$region)
```
part ii:
```{r}
# Create date_birth from yob, mob, dob
medrecord$date_birth<-make_date(day=medrecord$dob,month=medrecord$mob,year=medrecord$yob)
```
part iii:
```{r}
# Remove dob, mob, yob
medrecord_2<-medrecord%>%select(-dob,-mob,-yob)
```

Q2
part i:
```{r}
diagnoses<-read_csv("diagnoses.csv",show_col_types = FALSE)

# Convert date_diag to Date
diagnoses$date_diag<-as.Date(diagnoses$date_diag, format="%d/%m/%Y")
head(diagnoses$date_diag)
summary(diagnoses$date_diag)
```
part ii:
```{r}
# Label variables (optional with attributes)
attr(diagnoses,"label")<-"patient ID number"
attr(diagnoses,"label")<-"Date the diagnosis was made"
attr(diagnoses,"label")<-"Medical code assigned to diagnosis"
```
part iii:
```{r}
# Filter for Chikungunya fever (medcode 800)
diagnoses_cf<-diagnoses%>%filter(medcode==800)
view(diagnoses_cf)
```

Q3
part i:
```{r}
symptoms <- readRDS("Symptoms.RDS")
view(symptoms)

# Convert to long format
symp_long<-symptoms%>%pivot_longer(cols = -id,names_to = c(".value","visit"),names_pattern ="(.*)_(\\d+)")%>%mutate(visit=as.integer(visit))
view(symp_long)

# Create lowmood
symp_long<-symp_long%>%mutate(lowmood=if_else(mood<40,1,0))
view(symp_long)

# Identify and fix invalid fatigue
symp_long%>%filter(!fatigue%in%1:5)
symp_long<-symp_long%>%mutate(fatigue=if_else(fatigue>5,5,fatigue))
unique(symp_long$fatigue)
```
part ii:
```{r}
# Merge dataset and extra rows
str(medrecord_2$id)
str(diagnoses_cf$id)
medrecord_2<-medrecord_2%>%mutate(id=as.character(id))
diagnoses_cf<-diagnoses_cf%>%mutate(id=as.character(id))
merged_data<-left_join(medrecord_2,diagnoses_cf,by="id")
view(merged_data)

# Identify unmatched records
unmatched<-anti_join(diagnoses_cf,medrecord_2,by="id")
n_distinct(unmatched$id)

# Collapse to region and year
merged_data%>%mutate(year=year(date_diag))%>%group_by(region,year)%>%summarise(total_cases=n(),.groups = "drop")
```
a) Because some individuals were diagnosed with Chikungunya fever in both 2018 and 2019. Each diagnosis is as a separate row in the diagnoses_cf, the join results in duplicate IDs for people with two diagnoses, leading to more than 929 rows.
b)Yes, there are people in the diagnoses_cf who do not appear in the medical records medrecord_2.

part iii:
```{r}
medrecord_part2 <- readRDS("medrecord_part2.RDS")
view(medrecord_part2)

# Clean and align with medrecord_2
medrecord_part2 <- medrecord_part2%>%mutate(date_update=as.Date(substr(date_update, 5, nchar(date_update)),format="%d/%m/%Y"),region=as.factor(region),date_birth = dbirth)%>%select(-dbirth)
colnames(medrecord_part2)
colnames(medrecord_2)

# Bind together
medrecord_2<-medrecord_2%>%mutate(across(where(is.labelled),~as.characte))
medrecord_part2<-medrecord_part2%>%mutate(across(where(is.labelled),~as.character))
mr_sum<-bind_rows(medrecord_2,medrecord_part2)
```
part iv:
```{r}
# Merge with symp_long
symptom_merged<-inner_join(mr_sum, symp_long, by = "id")

# Summarise by fatigue
symptom_summary<-symptom_merged %>%group_by(fatigue) %>%summarise(num_visits = n(),visits_with_jointpain = sum(jointpain == 1),mean_mood = mean(mood),sd_mood = sd(mood), visits_with_lowmood =sum(lowmood == 1),.groups = "drop")
```

Q4
part i:
```{r}
treatments<-readRDS("treatment.RDS")
view(treatments)

# Calculate gap since last treatment
treatments<-treatment%>%arrange(id,date_treat)%>%group_by(id)%>%mutate(gap=as.numeric(date_treat-lag(date_treat)))%>%ungroup()
```
part ii:
```{r}
# First treatment
treat_f<-treatments%>%filter(!is.na(date_treat))%>%group_by(id)%>%slice_min(order_by = date_treat,n=1)%>%ungroup()
```
part iii:
```{r}
# Last treatment
treat_l<-treatments%>%group_by(id)%>%slice_max(order_by =date_treat,n=1,with_ties=FALSE)%>%ungroup()
```
